% -*- TeX -*- -*- UK -*- -*- Soft -*-

\chapter{DEAP}
\label{chap:DEAP}

This brief description is taken from \cite{DEAPGitHub2019}.

DEAP is a novel evolutionary computation framework for rapid prototyping and testing of ideas. It seeks to make algorithms explicit and data structures transparent. It works in perfect harmony with parallelisation mechanisms such as multiprocessing and SCOOP \cite{SCOOPDocs2019}.

DEAP includes the following features:
\begin{enumerate}
\item Genetic algorithm using any imaginable representation
\begin{enumerate}
\item List, Array, Set, Dictionary, Tree, Numpy Array, etc.
\end{enumerate}
\item Genetic programing using prefix trees
\begin{enumerate}
\item Loosely typed, Strongly typed
\item Automatically defined functions
\end{enumerate}
\item Evolution strategies (including CMA-ES)
\item Multi-objective optimisation (NSGA-II, SPEA2, MO-CMA-ES)
\item Co-evolution (cooperative and competitive) of multiple populations
\item Parallelization of the evaluations (and more)
\item Hall of Fame of the best individuals that lived in the population
\item Checkpoints that take snapshots of a system regularly
\item Benchmarks module containing most common test functions
\item Genealogy of an evolution (that is compatible with NetworkX)
\item Examples of alternative algorithms : Particle Swarm Optimization, Differential Evolution, Estimation of Distribution Algorithm

\end{enumerate}

Introduction

The DEAP framework [1; 2] is designed over the three following founding
principles:
\begin{enumerate}
\item 
Data structures are key to evolutionary computation. They must facilitate
the implementation of algorithms and be easy to customize.

\item 
Operator selection and algorithm parameters have strong influences
on evolutions, while often being problem dependent. Users should
be able to parametrize every aspect of the algorithms with minimal
complexity.

\item 
 EAs are usually embarrassingly parallel. Therefore, mechanisms
that implement distribution paradigms should be trivial to use.

\end{enumerate}

With the help of its sister project SCOOP [3] and the power of the Python
programming language, DEAP implements these three principles in a
simple and elegant design.


Data Structures

A very important part of the success for designing any algorithm - if
not the most important - is choosing the appropriate data structures.
Freedom in type creation is fundamental in the process of designing evolutionary
algorithms that solve real world problems. DEAP’s creator module
allows users to:
\begin{itemize}

\item 
create classes with a single line of code (inheritance);

\item 
add attributes (composition);

\item 
group classes in a single module (sandboxing).

\end{itemize}


\TBC{TBC}

